---
title: "Descriptives"
author: "Barry Sheppard"
date: "1 March 2015"
output: html_document
---

This document provides a breakdown of the statistical analysis from a research study.

Below loads the raw data file into the system and outputs the file list of headers in the data set.

```{r}
# Import dataset
dat <- read.csv("DisgustData.csv")
# Return the names of the datset
names(dat)
```


Next up the headers of the dataset are renamed to make things a little easier. 

```{r}
# Recoding the data for use

# Shorten some of the names
names(dat)[5] <- "Age"
names(dat)[6] <- "Sex"
names(dat)[7] <- "RWA01"
names(dat)[8] <- "RWA02"
names(dat)[9] <- "RWA03"
names(dat)[10] <- "RWA04"
names(dat)[11] <- "RWA05"
names(dat)[12] <- "RWA06"
names(dat)[13] <- "RWA07"
names(dat)[14] <- "RWA08"
names(dat)[15] <- "RWA09"
names(dat)[16] <- "RWA10"
names(dat)[17] <- "RWA11"
names(dat)[18] <- "RWA12"
names(dat)[19] <- "RWA13"
names(dat)[20] <- "RWA14"
names(dat)[21] <- "RWA15"
names(dat)[22] <- "RWA16"
names(dat)[23] <- "RWA17"
names(dat)[24] <- "RWA18"
names(dat)[25] <- "RWA19"
names(dat)[26] <- "RWA20"
names(dat)[27] <- "RWA21"
names(dat)[28] <- "RWA22"
names(dat)[29] <- "BeforeCricket"
names(dat)[30] <- "BeforeCricketBar"
names(dat)[31] <- "Group"
names(dat)[32] <- "AfterCricketBar"
names(dat)[33] <- "AfterCricket"
names(dat)[34] <- "Comments"

```


The Right Wing Authoritarianism scale still needs to be calculated. 

We also need to calculate the changes in rating for the food types and add these to the dataset.



```{r}

# Function to change the Likert results from the long form to a number
library(plyr)
likert_to_number <- function(x){
  x <- revalue(x, c("-4 You very strongly disagree with the statement."="-4",
                                    "-3 You strongly disagree with the statement. "="-3",
                                    "-2 You moderately disagree with the statement."="-2",
                                    "-1 You slightly disagree with the statement."="-1",
                                    "0 You feel exactly and precisely neutral about an item."="0",
                                    "+1 You slightly agree with the statement. "="1",
                                    "+2 You moderately agree with the statement."="2",
                                    "+3 You strongly agree with the statement. "="3",
                                    "+4 You very strongly agree with the statement."="4"
 )   )
  
  return(x)
}

dat$RWA01 <- likert_to_number(dat$RWA01)
dat$RWA02 <- likert_to_number(dat$RWA02)
dat$RWA03 <- likert_to_number(dat$RWA03)
dat$RWA04 <- likert_to_number(dat$RWA04)
dat$RWA05 <- likert_to_number(dat$RWA05)
dat$RWA06 <- likert_to_number(dat$RWA06)
dat$RWA07 <- likert_to_number(dat$RWA07)
dat$RWA08 <- likert_to_number(dat$RWA08)
dat$RWA09 <- likert_to_number(dat$RWA09)
dat$RWA10 <- likert_to_number(dat$RWA10)
dat$RWA11 <- likert_to_number(dat$RWA11)
dat$RWA12 <- likert_to_number(dat$RWA12)
dat$RWA13 <- likert_to_number(dat$RWA13)
dat$RWA14 <- likert_to_number(dat$RWA14)
dat$RWA15 <- likert_to_number(dat$RWA15)
dat$RWA16 <- likert_to_number(dat$RWA16)
dat$RWA17 <- likert_to_number(dat$RWA17)
dat$RWA18 <- likert_to_number(dat$RWA18)
dat$RWA19 <- likert_to_number(dat$RWA19)
dat$RWA20 <- likert_to_number(dat$RWA20)
dat$RWA21 <- likert_to_number(dat$RWA21)
dat$RWA22 <- likert_to_number(dat$RWA22)


# Next lets score the RWA results

# Scores go from 20 to 180
# Question 1 and 2 are test questions and not used
# Positive scores are 3, 5, 7, 10, 12, 14, 16, 17, 19 and 22 
# Add all the positive scores and add 50
# The remainder are reversed questions, we could recode but its easier to do the following
# Start with 50 and substract the total of the negative scores

dat$RWAscore <- 100 + as.numeric(as.character(dat$RWA03)) + as.numeric(as.character(dat$RWA05)) + 
  as.numeric(as.character(dat$RWA07)) + as.numeric(as.character(dat$RWA10)) + as.numeric(as.character(dat$RWA12)) + 
  as.numeric(as.character(dat$RWA14)) + as.numeric(as.character(dat$RWA16)) + as.numeric(as.character(dat$RWA17)) + 
  as.numeric(as.character(dat$RWA19)) + as.numeric(as.character(dat$RWA22)) - as.numeric(as.character(dat$RWA04)) - 
  as.numeric(as.character(dat$RWA06)) - as.numeric(as.character(dat$RWA08)) - as.numeric(as.character(dat$RWA09)) - 
  as.numeric(as.character(dat$RWA11)) - as.numeric(as.character(dat$RWA13)) - as.numeric(as.character(dat$RWA15)) - 
  as.numeric(as.character(dat$RWA18)) - as.numeric(as.character(dat$RWA20)) - as.numeric(as.character(dat$RWA21))

# Calculate the changes in rating
dat$CricketChange <- dat$AfterCricket - dat$BeforeCricket
dat$BarChange <- dat$AfterCricketBar - dat$BeforeCricket

# Return the updated names of the dataset
names(dat)

```


Some basic descriptive info on the data in numbers

```{r}

library(psych)
describe(dat$Age)
describe(dat$BeforeCricket)
describe(dat$BeforeCricketBar)
describe(dat$AfterCricket)
describe(dat$AfterCricketBar)
describe(dat$RWAscore)
describe(dat$CricketChange)
describe(dat$BarChange)


```

Lets produce some charts to visualise the data

[Descriptives 02](/descriptives02.html)


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plots.
